---
title: "Prioritization of cropland restoration"
authors: "Nickolas McManus and Michelle Geldin"
date: '2022-10-13'
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(prioritizr)    # prioritization package 
library(gurobi)        # optimizaiton software
library(sf)            # associated with prioritizr
library(terra)         # for reading in and altering rasters
library(raster)        # for formating rasters in prioritizr-friendly format
library(geobr)         # geometries for Brazilian boundaries
library(ggplot2)       # visualize some outputs
library(tidyverse)     # always!
library(janitor)       # cleaning up dfs
library(kableExtra)    # tables
```

## READ IN & CROP DATA

First we will read in all the data necessary for formulating and solving the prioritization problem. Broadly, this includes the planning units (locations of projected cropland abandonment), the features (benefits to biodiversity and carbon), and costs of restoration. Additionally, we need to read in data regarding Brazilian administrative and ecological boundaries. These boundaries are used to crop planning unit and feature layers and ensure matching spatial extents and resolutions.


### Cropland Abandonment

Read in areas of projected cropland abandonment for five different SSP scenarios in Brazil. These rasters were generated from code located in the `abandoned-cropland` repository, which provides a detailed methodology. Pixel values range from 1 to 7 indicating the frequency of abandonment between 2020 and 2050; values of 1 indicate abandonment only occurred for a five year duration, while pixels with values of 7 were projected to be abandoned for all of 2020-2050.  

Projections of land-use under different SSP scenarios were sourced from:
- Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6

```{r}
# read in the cropland abandonment
ssp1_abandoned_crop <- rast("data/inputs/ssp1/SSP1_abandoned_cropland_brazil.tif")
ssp2_abandoned_crop <- rast("data/inputs/ssp2/SSP2_abandoned_cropland_brazil.tif")
ssp3_abandoned_crop <- rast("data/inputs/ssp3/SSP3_abandoned_cropland_brazil.tif")
ssp4_abandoned_crop <- rast("data/inputs/ssp4/SSP4_abandoned_cropland_brazil.tif")
ssp5_abandoned_crop <- rast("data/inputs/ssp5/SSP5_abandoned_cropland_brazil.tif")
```

We also want to figure out which parcels of abandoned cropland are present in every SSP scenario. To solve this, we'll first reclassify the abandoned crop layers to just be a value of 1 (rather than range of 1-7). Then, we'll add these into the combined raster. 
```{r}
reclass_df <- c(1, 7, 1)
reclass_m <- matrix(reclass_df, ncol = 3, byrow = T)

ssp1_reclass <- classify(ssp1_abandoned_crop, reclass_m)
ssp2_reclass <- classify(ssp2_abandoned_crop, reclass_m)
ssp3_reclass <- classify(ssp3_abandoned_crop, reclass_m)
ssp4_reclass <- classify(ssp4_abandoned_crop, reclass_m)
ssp5_reclass <- classify(ssp5_abandoned_crop, reclass_m)

# add together and only keep planning units that overlap in all SSPs
ssp_all_abandoned_crop <- ssp1_reclass + ssp2_reclass + ssp3_reclass + ssp4_reclass + ssp5_reclass
```



### Brazilian Biomes

The boundary information for Brazilian biomes is read-in using the geobr package. Biome polygons are sourced from the Brazilian Institute of Geography and Statistics (IBGE). This dataset uses 2019 IBGE data at scale 1:250.000. Because carbon data is missing for the majority of the pantanal biome, this region is removed from the remaining planning unit and feature layers. 

Source: https://www.ibge.gov.br/geociencias/cartas-e-mapas/informacoes-ambientais/15842-biomas.html?=&t=acesso-ao-produto

```{r}
# read in biome boundaries from geobr package
# NOTE: this may take a while to locally download
biomes <- read_biomes(
  year = 2019,         #the most recent data available
  simplified = FALSE   #full resolution for boundaries
)

# create vector for whole of Brazil
biomes_vect <- biomes[-7, ] %>%  #remove coastal system biome
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% #change crs to match other rasters
  vect()  #turn sf into vector; easier to crop rasts with


# create vectors for each of the biomes ---------------------------------------
amazon <- biomes %>% 
  filter(code_biome == 1) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

caatinga <- biomes %>% 
  filter(code_biome == 2) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

cerrado <- biomes %>% 
  filter(code_biome == 3) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

atlantic_forest <- biomes %>% 
  filter(code_biome == 4) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pampa <- biomes %>% 
  filter(code_biome == 5) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pantanal <- biomes %>% 
  filter(code_biome == 6) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()
```



### Carbon

This dataset, originally created by Cook-Patton et al. and updated by Global Forest Watch, estimates the rate of carbon sequestration in aboveground and belowground live biomass during the first 30 years of natural forest regeneration. Spatial sequestration estimates include all forest and savanna biomes in units of MgC/ha/yr at a ~1 km resolution. This carbon data is cropped, masked, and reprojected to match the spatial extent and resolution of other rasters used in the prioritization. Because sequestration data is mostly missing for the Pantanal biome in Brazil, it will be removed from this and all data in the analysis. 

Source: https://data.globalforestwatch.org/documents/gfw::carbon-accumulation-potential-from-natural-forest-regrowth-in-forest-and-savanna-biomes/about

```{r, eval=FALSE}
#NOTE: this code chunk defaults to not evaluate.
#only needs to be run once and cropped raster is saved

# read in carbon layer
carbon_global <- rast("data/inputs/sequestration_rate_mean_aboveground_full_extent_Mg_C_ha_yr.tif")

#cannot crop/mask carbon data when biomes are different crs
#change biome crs because it's faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(carbon_global))

# crop carbon data to Brazil
carbon_brazil <- carbon_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>%  
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect)) 

# default resolution of carbon layer is 898.8m
# change resolution to match other rasters (1km)
carbon_brazil <- resample(carbon_brazil, ssp1_abandoned_crop, method = "bilinear")

# remove Panatnal biome from carbon layer
carbon_brazil <- mask(carbon_brazil, pantanal, inverse = TRUE, updatevalue = NA, touches = TRUE)

# change the name saved within raster for easier recall
# after stacking feature rasters
names(carbon_brazil) <- "carbon_brazil"

# test plot to visualize results
plot(carbon_brazil)
plot(biomes_vect, add = TRUE)

#export carbon_brazil as intermediate raster
writeRaster(carbon_brazil, "data/outputs/carbon_seq_rate_Brazil.tif", overwrite = TRUE)
```

```{r}
# run below line of code just to read in layer if already created
carbon_brazil <- rast("data/outputs/carbon_seq_rate_Brazil.tif")
```



### Biodiversity

After reading in the global layer, we will crop it to Brazil and resample to match the resolution of the planning units. Then, we will reclassify the layer to extinction risk values; this will help set tangible goals of minimizing extinction risk. The final raster will be saved locally, allowing the user to simply read in the layer in the future and only run this code-chunk once. 

```{r, eval=FALSE}
###NOTE: this code chunk defaults to not evaluate.
###only needs to be run once and then cropped raster is saved

#read in global biodiversity data
biodiversity_global <- rast("data/inputs/sparc_conservationPriorities.tif")


#temporarily change biome crs to crop/mask biodiversity data
#faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(biodiversity_global))


# crop biodiversity layer to Brazil
biodiversity_brazil <- biodiversity_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect))


# change resolution of biodiversity layer (4.72km) to match other rasters (1km)
# first use terra:disagg()
biodiversity_brazil <- disagg(biodiversity_brazil, 
                              fact = 5,              
                              method = "bilinear")  
# resample higher res biodiversity to ensure same res and geometries 
# (fine-tuning the res)
biodiversity_brazil <- resample(biodiversity_brazil,ssp1_abandoned_crop, method = "bilinear")

# remove Pantanal biome from biodiversity layer
biodiversity_brazil <- mask(biodiversity_brazil, pantanal, inverse = TRUE, updatevalue = NA)

# change the name saved within raster for easier recall
# after stacking feature rasters
names(biodiversity_brazil) <- "biodiversity_brazil"

# test plot to visualize results
plot(biodiversity_brazil)
plot(biomes_vect, add = TRUE)

# export biodiversity_brazil as intermediate raster
writeRaster(biodiversity_brazil, 
            "data/outputs/sparc_conservationPriorities_Brazil.tif", 
            overwrite = TRUE)




### Now reclassify cropped layer to extinction risk -----------------------------------------
## read in values
extrisk <- read_csv("data/inputs/globalverts_extrisk_scale.csv")

## make three columns for a "from to" interval reclass
extrisk_interval <- extrisk %>%
  # bring rasterval from previous row down as lower bound
  mutate(from = lag(rasterval), .before = 1) %>% 
  # rename rasterval to indicate upper bound
  rename(to = rasterval) %>% 
  # remove resulting na
  drop_na 

# turn csv into matrix
extrisk_m <- data.matrix(extrisk_interval, rownames.force = NA)

# reclassify raster to extrisk
bio_ext_brazil <- classify(biodiversity_brazil, extrisk_m, right = TRUE)

## save new raster
writeRaster(bio_ext_brazil, "data/outputs/biodiversity_extrisk_brazil.tif")
```

Run below line of code just to read in layer if already created:
```{r}
bio_ext_brazil <- rast("data/outputs/biodiversity_extrisk_brazil.tif")
```



### Cost Layer

The cost of restoration in Brazil differs with location (by biome), restoration method (six types), and environmental condition of land (favorable or unfavorable). These costs were sourced from a 2017 Nature Conservancy report, written by Rubens de Miranda Benini and Sergio Adeodato. 
```{r}
#read in cost data
costs <- read_csv("data/inputs/tnc_cost_biomes_methods.csv") %>% 
  clean_names()
```







## PREP DATA FOR PRIORITIZATION

This section wrangles data into formats suitable for `prioritizr` inputs. 

### Cost

Select the restoration method and environmental condition you want to investigate for the cost layer. Enter these in the `filter()` lines of code. 
Costs are per hectare, so we need to convert it to km^2. This makes the cost numbers too large for `prioritizr`, so we need to convert costs to millions per km^2. 
```{r}
# Looking at conducting natural regeneration and "unfavorable environmental conditions" (CAD)
costs_filtered <- costs %>% 
  filter(restoration_method == "conducting_natural_regeneration") %>% 
  filter(environmental_conditions == "CAD") %>% 
  filter(biome_code != 6) %>%   #remove pantanal data
  select(!c(id, biome)) %>%   #remove id and biome columns
  mutate(cost_mil_r_km = cost_r_ha / 10000,   #convert to cost/km2
         cost_mil_usd_km = cost_usd_ha / 10000,
         .before = 1)

```

To properly assign costs to our planning units, we need to merge our biome data with the filtered cost data. This dataset will then be rasterized for the next step of finalizing our planning units. 
```{r}
# first join biomes with cost layer 
biomes_cost <- biomes %>% 
  full_join(costs_filtered, by = c("code_biome" = "biome_code"))

# turn cost/biome data into a vector
biomes_cost_vect <- biomes_cost[-7, ] %>%   #remove coastal system biome
  st_transform(crs = crs(ssp1_abandoned_crop)) %>%  #change crs to match other rasters
  vect()  

# turn vect into a raster
biomes_cost_rast <- rasterize(x = biomes_cost_vect,  #vect file to rasterize
                              #raster file to provide example crs and spatial extent
                              y = ssp1_abandoned_crop,  
                              #values for raster
                              field = "cost_mil_r_km")      

# remove pantanal biome
biomes_cost_rast <- mask(biomes_cost_rast, pantanal, inverse = TRUE, updatevalue = NA)
```


### Planning Units

After deciding the restoration method and land status for cost (previous section), assign costs to each planning unit.
```{r}
# mask the cost by biomes to abandoned cropland raster
# this will isolate pixels of interest (planning units) with associated cost value
ssp1_pu <- mask(biomes_cost_rast, ssp1_abandoned_crop)
ssp2_pu <- mask(biomes_cost_rast, ssp2_abandoned_crop)
ssp3_pu <- mask(biomes_cost_rast, ssp3_abandoned_crop)
ssp4_pu <- mask(biomes_cost_rast, ssp4_abandoned_crop)
ssp5_pu <- mask(biomes_cost_rast, ssp5_abandoned_crop)
ssp_all_pu <- mask(biomes_cost_rast, ssp_all_abandoned_crop)
```

`prioritizr` requires rasters to be in "RasterLayer" format, which requires converting rasters from `terra` package format to the `raster` package format. 

**NOTE:** Often this conversion throws an error when first run (cannot determine the reason). The second time it's run, it may convert raster formats but then drops all the values. Often you have to run this code chunk 2-3 times before obtaining properly converting RasterLayers for all the planning units. Occasionally this involves re-runing the previous code chunk as well. Double-check that they have converted properly before moving on!
```{r}
# double-check values before proceeding!!
# often need to run this 3 times for correct conversion
ssp1_pu <- raster(ssp1_pu)
ssp2_pu <- raster(ssp2_pu)
ssp3_pu <- raster(ssp3_pu)
ssp4_pu <- raster(ssp4_pu)
ssp5_pu <- raster(ssp5_pu)
ssp_all_pu <- raster(ssp_all_pu)
```



### Features

To evaluate carbon and biodiversity as features in the prioritization problem, they must be merged into a raster stack.

Here we will mask the biodiversity and carbon layers to the planning units available for in each SSP. This will result in feature layers that reflect the total amount that could be achieved by restoring every planning unit. This makes it easier to set relative targets based on what is possible in the prioritization. However, it means creating a different feature stack for each SSP. Below we will create a function to generate the necessary feature data rasters, convert them to a prioritizr-friendly raster format, and then stack them according to SSP. 

**NOTE:** As in the previous section, converting raster formats and stacking sometimes returns an error. Double-check the values of the raster stacks before proceeding. This code chunk may need to be run 2-3 times. 
```{r}
## function will mask Brazil-wide carbon and biodiversity layers
## by planning units for each SSP, then stack raster 
mask_feat_pu <- function(carbon, bio, ssp) {
  # mask brazil carbon layer with SSP pus, then convert to rastLayer
  carbon_pu_ssp <- terra::mask(carbon, ssp) 
  carbon_pu_ssp <- raster(carbon_pu_ssp)
  
  # mask biodiversity layer by SSP pus, then convert
  bio_pu_ssp <- terra::mask(bio, ssp)
  bio_pu_ssp <- raster(bio_pu_ssp)
  
  # stack the rastLayers
  feature_pu_ssp <- stack(carbon_pu_ssp, bio_pu_ssp)
  return(feature_pu_ssp)
}

## Now use function to create feature rastStack for each SSP
## (in future, figure out how to loop this!)
ssp1_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp1_abandoned_crop)
ssp2_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp2_abandoned_crop)
ssp3_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp3_abandoned_crop)
ssp4_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp4_abandoned_crop)
ssp5_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp5_abandoned_crop)
ssp_all_feat <- mask_feat_pu(carbon_brazil, bio_ext_brazil, ssp_all_abandoned_crop)
```





## PROBLEM FORMULATION

This is where we will start formulating the problem to be solved in prioritizr! A base problem will be created for each SSP using their specific planning units and feature layers. All base problems aim to minimize target shortfall while staying within a budget. Relative targets are set for carbon and biodiversity, higher than is possible within the budget (50%). In the spatial solution, it will attempt to get as close to this unattainable target as possible while staying within the budget. These relative targets can be set higher (e.g. 100%), but due to increased shortfall and the pressure to minimize it, results may skew more heavily to certain biomes with lower restoration costs. The model sensitivity to relative targets should be considered when using this general methodology for other features or geographic areas. We recommend first ascertaining the relative percent of features that can realistically be restored under a given budget, and using this value to set the relative targets under a minimized shortfall objective. 


```{r}
p1 <- problem(ssp1_pu, ssp1_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)

p2 <- problem(ssp2_pu, ssp2_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)

p3 <- problem(ssp3_pu, ssp3_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)

p4 <- problem(ssp4_pu, ssp4_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)

p5 <- problem(ssp5_pu, ssp5_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)

p_all <- problem(ssp_all_pu, ssp_all_feat) %>% 
  add_relative_targets(c(0.5, 0.5)) %>% 
  add_gurobi_solver(gap = 0.05)
```


### Report outputs

We can use the base problems to evaluate solutions based on more customizable parameters. 
For example, in this analysis we evaluated difference in solutions based on available budget, while maintaining and equal weight for the feature layer.

The low-end budget is 445.892 million Brazilian Reals. To calculate this amount, we looked at the Environmental management budget of Brazil's Ministry of the Environment. In 2022, Brazil only used 85.8% of the allotted budget. We're curious how much land could be restored with the budget "leftovers". This percentage was used to predict how much money may go unused within the Environmental management budget for 2023: roughly BRL445,862,335. 

To determine the potential high-end of a budget, we looked toward the Amazon Fund. This fund receives money from organizations and countries around the world and can be used at the discretion of the Brazilian government to conserve and restore land throughout the country. NGOs can utilize this fund for projects within the Amazon biome, but the Federal government can use these funds for projects throughout the country. We'll evaluate how much abandoned cropland could be restored if Brazil utilized the entire budget currently available in the Amazon fund: BRL3,400,000,000. As a reminder, costs are converted to units of "millions" so that prioritizr can run correctly.

Solutions are output as a rasterstack for each SSP: one layer representing low budget solutions and another layer with the high budget solutions.

**NOTE:** For ease this was done by creating a function and applying to all SSP base problems. The down-side is that you cannot easily use prioritizr evaluation functions such as `eval_target_coverage_summary()` because the budget-based problem doesn't exist as an object outside of the function. If you'd like to evaluate summary statistics using these functions in the prioritizr package, there is commented-out example code for how to run this manually outside of the function at the bottom of the code chunk. 

```{r}
### Create function --------------------------------------
## Takes base problem, low and high budgets and then returns raster stack.
## Uses default values for budgets. Can be changed by user when calling fxn 
budget_solutions <- function(base_p, low_bud = 445.862, high_bud = 3400) {
  ## Create problems for low and high budget
  p_lowBud <- base_p %>% 
    add_min_shortfall_objective(budget = low_bud)
  
  p_highBud <- base_p %>% 
    add_min_shortfall_objective(budget = high_bud)
  
  ## Solve in a stack
  s <- stack(solve(p_lowBud), solve(p_highBud))
  
  ## rename rasters in stack for clarity
  names(s)[1] = "lowBud"
  names(s)[2] = "highBud"
  
  ## return the stacked solution
  return(s)
}


### Run fxn across all SSP base problems ---------------

## SSP1
s1 <- budget_solutions(p1)
## SSP2
s2 <- budget_solutions(p2)
## SSP3
s3 <- budget_solutions(p3)
## SSP4
s4 <- budget_solutions(p4)
## SSP5
s5 <- budget_solutions(p5)
## SSP Overlap
s_all <- budget_solutions(p_all)

### Export solutions ------------------------------------
## convert RasterLayers solutions back to SpatRast and assign crs
ssp1_sol <- rast(s1)
crs(ssp1_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp1_sol, "data/outputs/ssp1_solution_country.tif", overwrite = TRUE)

ssp2_sol <- rast(s2)
crs(ssp2_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp2_sol, "data/outputs/ssp2_solution_country.tif", overwrite = TRUE)

ssp3_sol <- rast(s3)
crs(ssp3_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp3_sol, "data/outputs/ssp3_solution_country.tif", overwrite = TRUE)

ssp4_sol <- rast(s4)
crs(ssp4_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp4_sol, "data/outputs/ssp4_solution_country.tif", overwrite = TRUE)

ssp5_sol <- rast(s5)
crs(ssp5_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp5_sol, "data/outputs/ssp5_solution_country.tif", overwrite = TRUE)

ssp_all_sol <- rast(s_all)
crs(ssp_all_sol) = crs(ssp1_abandoned_crop)
writeRaster(ssp_all_sol, "data/outputs/ssp_all_solution_country.tif", overwrite = TRUE)


# ### SSP1 -----------------------------------------------------
# 
# ## create problems --------------
# # low budget
# p1_lowBud <- p1 %>% 
#   add_min_shortfall_objective(budget = 445.862) 
# 
# # high budget
# p1_highBud <- p1 %>% 
#   add_min_shortfall_objective(budget = 3400) 
# 
# ## solve --------------------------
# s1 <- stack(solve(p1_lowBud), solve(p1_highBud))
# # rename rasters in stack
# names(s1)[1] = "lowBud"
# names(s1)[2] = "highBud"

# ## get solution stats --------------
# # returns the number of pus chosen in solution for low budget
# eval_target_coverage_summary(p1_lowBud, s1$lowBud)
# # returns amount of features held in solution for low budget
# eval_target_coverage_summary(p1_lowBud, s1$lowBud)
```
 


### Shiny App outputs

For greater user flexibility in our Shiny App (linked in repo README file), we've also generated prioritization results based on different feature weights. These range from weighing biodiversity 5 times greater than carbon, to weighing carbon 5 times greater than biodiversity. As a result, each SSP has 5 solution outputs for a given budget scenario. With a low and high-end budget, each SSP has a total of 10 solutions. These solution rasters are exported to their own directory and utilized for visualization and calculating zonal statistics in the interactive Shiny App. 

```{r}
## Function to find solutions based on feature weights 
feat_weights <- function(weights, bud, base_p) {
  p <- base_p %>% 
    add_min_shortfall_objective(budget = bud) %>% 
    add_feature_weights(weights)
  
  s <- solve(p)
  return(s)
}

## SSP1 ------------------------------------------------------------------------

## Low budget solutions
ssp1_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p1)
ssp1_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p1)
ssp1_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p1)
ssp1_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p1)
ssp1_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p1)

# rename rasters
names(ssp1_lowBud_15) = "ssp1_lowBud_feat_15"
names(ssp1_lowBud_24) = "ssp1_lowBud_feat_24"
names(ssp1_lowBud_33) = "ssp1_lowBud_feat_33"
names(ssp1_lowBud_42) = "ssp1_lowBud_feat_42"
names(ssp1_lowBud_51) = "ssp1_lowBud_feat_51"

# Export
ssp1_lowBud_15_rast <- rast(ssp1_lowBud_15)
crs(ssp1_lowBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_lowBud_15_rast, "data/outputs/shiny/ssp1_lowBud_feat15.tif", overwrite = T)

ssp1_lowBud_24_rast <- rast(ssp1_lowBud_24)
crs(ssp1_lowBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_lowBud_24_rast, "data/outputs/shiny/ssp1_lowBud_feat24.tif", overwrite = T)

ssp1_lowBud_33_rast <- rast(ssp1_lowBud_33)
crs(ssp1_lowBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_lowBud_33_rast, "data/outputs/shiny/ssp1_lowBud_feat33.tif", overwrite = T)

ssp1_lowBud_42_rast <- rast(ssp1_lowBud_42)
crs(ssp1_lowBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_lowBud_42_rast, "data/outputs/shiny/ssp1_lowBud_feat42.tif", overwrite = T)

ssp1_lowBud_51_rast <- rast(ssp1_lowBud_51)
crs(ssp1_lowBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_lowBud_51_rast, "data/outputs/shiny/ssp1_lowBud_feat51.tif", overwrite = T)


## High budget
ssp1_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p1)
ssp1_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p1)
ssp1_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p1)
ssp1_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p1)
ssp1_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p1)

# rename rasters
names(ssp1_highBud_15) = "ssp1_highBud_feat15"
names(ssp1_highBud_24) = "ssp1_highBud_feat24"
names(ssp1_highBud_33) = "ssp1_highBud_feat33"
names(ssp1_highBud_42) = "ssp1_highBud_feat42"
names(ssp1_highBud_51) = "ssp1_highBud_feat51"

# Export
ssp1_highBud_15_rast <- rast(ssp1_highBud_15)
crs(ssp1_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_highBud_15_rast, "data/outputs/shiny/ssp1_highBud_feat15.tif", overwrite = T)

ssp1_highBud_24_rast <- rast(ssp1_highBud_24)
crs(ssp1_highBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_highBud_24_rast, "data/outputs/shiny/ssp1_highBud_feat24.tif", overwrite = T)

ssp1_highBud_33_rast <- rast(ssp1_highBud_33)
crs(ssp1_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_highBud_33_rast, "data/outputs/shiny/ssp1_highBud_feat33.tif", overwrite = T)

ssp1_highBud_42_rast <- rast(ssp1_highBud_42)
crs(ssp1_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_highBud_42_rast, "data/outputs/shiny/ssp1_highBud_feat42.tif", overwrite = T)

ssp1_highBud_51_rast <- rast(ssp1_highBud_51)
crs(ssp1_highBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp1_highBud_51_rast, "data/outputs/shiny/ssp1_highBud_feat51.tif", overwrite = T)




## SSP2 ------------------------------------------------------------------------

## Low budget 
ssp2_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p2)
ssp2_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p2)
ssp2_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p2)
ssp2_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p2)
ssp2_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p2)

# rename rasters
names(ssp2_lowBud_15) = "ssp2_lowBud_feat_15"
names(ssp2_lowBud_24) = "ssp2_lowBud_feat_24"
names(ssp2_lowBud_33) = "ssp2_lowBud_feat_33"
names(ssp2_lowBud_42) = "ssp2_lowBud_feat_42"
names(ssp2_lowBud_51) = "ssp2_lowBud_feat_51"

# Export
ssp2_lowBud_15_rast <- rast(ssp2_lowBud_15)
crs(ssp2_lowBud_15_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_lowBud_15_rast, "data/outputs/shiny/ssp2_lowBud_feat15.tif", overwrite = T)

ssp2_lowBud_24_rast <- rast(ssp2_lowBud_24)
crs(ssp2_lowBud_24_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_lowBud_24_rast, "data/outputs/shiny/ssp2_lowBud_feat24.tif", overwrite = T)

ssp2_lowBud_33_rast <- rast(ssp2_lowBud_33)
crs(ssp2_lowBud_33_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_lowBud_33_rast, "data/outputs/shiny/ssp2_lowBud_feat33.tif", overwrite = T)

ssp2_lowBud_42_rast <- rast(ssp2_lowBud_42)
crs(ssp2_lowBud_42_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_lowBud_42_rast, "data/outputs/shiny/ssp2_lowBud_feat42.tif", overwrite = T)

ssp2_lowBud_51_rast <- rast(ssp2_lowBud_51)
crs(ssp2_lowBud_51_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_lowBud_51_rast, "data/outputs/shiny/ssp2_lowBud_feat51.tif", overwrite = T)

## High budget
ssp2_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p2)
ssp2_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p2)
ssp2_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p2)
ssp2_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p2)
ssp2_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p2)

# rename rasters
names(ssp2_highBud_15) = "ssp2_highBud_feat15"
names(ssp2_highBud_24) = "ssp2_highBud_feat24"
names(ssp2_highBud_33) = "ssp2_highBud_feat33"
names(ssp2_highBud_42) = "ssp2_highBud_feat42"
names(ssp2_highBud_51) = "ssp2_highBud_feat51"

# Export
ssp2_highBud_15_rast <- rast(ssp2_highBud_15)
crs(ssp2_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp2_highBud_15_rast, "data/outputs/shiny/ssp2_highBud_feat15.tif", overwrite = T)

ssp2_highBud_24_rast <- rast(ssp2_highBud_24)
crs(ssp2_highBud_24_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_highBud_24_rast, "data/outputs/shiny/ssp2_highBud_feat24.tif", overwrite = T)

ssp2_highBud_33_rast <- rast(ssp2_highBud_33)
crs(ssp2_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp2_highBud_33_rast, "data/outputs/shiny/ssp2_highBud_feat33.tif", overwrite = T)

ssp2_highBud_42_rast <- rast(ssp2_highBud_42)
crs(ssp2_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp2_highBud_42_rast, "data/outputs/shiny/ssp2_highBud_feat42.tif", overwrite = T)

ssp2_highBud_51_rast <- rast(ssp2_highBud_51)
crs(ssp2_highBud_51_rast) <- crs(ssp2_abandoned_crop)
writeRaster(ssp2_highBud_51_rast, "data/outputs/shiny/ssp2_highBud_feat51.tif", overwrite = T)




## SSP3 ------------------------------------------------------------------------

## Low budget 
ssp3_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p3)
ssp3_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p3)
ssp3_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p3)
ssp3_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p3)
ssp3_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p3)

# rename rasters
names(ssp3_lowBud_15) = "ssp3_lowBud_feat_15"
names(ssp3_lowBud_24) = "ssp3_lowBud_feat_24"
names(ssp3_lowBud_33) = "ssp3_lowBud_feat_33"
names(ssp3_lowBud_42) = "ssp3_lowBud_feat_42"
names(ssp3_lowBud_51) = "ssp3_lowBud_feat_51"

# Export
ssp3_lowBud_15_rast <- rast(ssp3_lowBud_15)
crs(ssp3_lowBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_lowBud_15_rast, "data/outputs/shiny/ssp3_lowBud_feat15.tif", overwrite = T)

ssp3_lowBud_24_rast <- rast(ssp3_lowBud_24)
crs(ssp3_lowBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_lowBud_24_rast, "data/outputs/shiny/ssp3_lowBud_feat24.tif", overwrite = T)

ssp3_lowBud_33_rast <- rast(ssp3_lowBud_33)
crs(ssp3_lowBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_lowBud_33_rast, "data/outputs/shiny/ssp3_lowBud_feat33.tif", overwrite = T)

ssp3_lowBud_42_rast <- rast(ssp3_lowBud_42)
crs(ssp3_lowBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_lowBud_42_rast, "data/outputs/shiny/ssp3_lowBud_feat42.tif", overwrite = T)

ssp3_lowBud_51_rast <- rast(ssp3_lowBud_51)
crs(ssp3_lowBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_lowBud_51_rast, "data/outputs/shiny/ssp3_lowBud_feat51.tif", overwrite = T)


## High budget
ssp3_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p3)
ssp3_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p3)
ssp3_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p3)
ssp3_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p3)
ssp3_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p3)

# rename rasters
names(ssp3_highBud_15) = "ssp3_highBud_feat15"
names(ssp3_highBud_24) = "ssp3_highBud_feat24"
names(ssp3_highBud_33) = "ssp3_highBud_feat33"
names(ssp3_highBud_42) = "ssp3_highBud_feat42"
names(ssp3_highBud_51) = "ssp3_highBud_feat51"

# Export
ssp3_highBud_15_rast <- rast(ssp3_highBud_15)
crs(ssp3_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_highBud_15_rast, "data/outputs/shiny/ssp3_highBud_feat15.tif", overwrite = T)

ssp3_highBud_24_rast <- rast(ssp3_highBud_24)
crs(ssp3_highBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_highBud_24_rast, "data/outputs/shiny/ssp3_highBud_feat24.tif", overwrite = T)

ssp3_highBud_33_rast <- rast(ssp3_highBud_33)
crs(ssp3_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_highBud_33_rast, "data/outputs/shiny/ssp3_highBud_feat33.tif", overwrite = T)

ssp3_highBud_42_rast <- rast(ssp3_highBud_42)
crs(ssp3_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_highBud_42_rast, "data/outputs/shiny/ssp3_highBud_feat42.tif", overwrite = T)

ssp3_highBud_51_rast <- rast(ssp3_highBud_51)
crs(ssp3_highBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp3_highBud_51_rast, "data/outputs/shiny/ssp3_highBud_feat51.tif", overwrite = T)



## SSP4 ------------------------------------------------------------------------

## Low budget 
ssp4_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p4)
ssp4_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p4)
ssp4_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p4)
ssp4_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p4)
ssp4_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p4)

# rename rasters
names(ssp4_lowBud_15) = "ssp4_lowBud_feat_15"
names(ssp4_lowBud_24) = "ssp4_lowBud_feat_24"
names(ssp4_lowBud_33) = "ssp4_lowBud_feat_33"
names(ssp4_lowBud_42) = "ssp4_lowBud_feat_42"
names(ssp4_lowBud_51) = "ssp4_lowBud_feat_51"

# Export
ssp4_lowBud_15_rast <- rast(ssp4_lowBud_15)
crs(ssp4_lowBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_lowBud_15_rast, "data/outputs/shiny/ssp4_lowBud_feat15.tif", overwrite = T)

ssp4_lowBud_24_rast <- rast(ssp4_lowBud_24)
crs(ssp4_lowBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_lowBud_24_rast, "data/outputs/shiny/ssp4_lowBud_feat24.tif", overwrite = T)

ssp4_lowBud_33_rast <- rast(ssp4_lowBud_33)
crs(ssp4_lowBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_lowBud_33_rast, "data/outputs/shiny/ssp4_lowBud_feat33.tif", overwrite = T)

ssp4_lowBud_42_rast <- rast(ssp4_lowBud_42)
crs(ssp4_lowBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_lowBud_42_rast, "data/outputs/shiny/ssp4_lowBud_feat42.tif", overwrite = T)

ssp4_lowBud_51_rast <- rast(ssp4_lowBud_51)
crs(ssp4_lowBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_lowBud_51_rast, "data/outputs/shiny/ssp4_lowBud_feat51.tif", overwrite = T)


## High budget
ssp4_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p4)
ssp4_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p4)
ssp4_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p4)
ssp4_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p4)
ssp4_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p4)

# rename rasters
names(ssp4_highBud_15) = "ssp4_highBud_feat15"
names(ssp4_highBud_24) = "ssp4_highBud_feat24"
names(ssp4_highBud_33) = "ssp4_highBud_feat33"
names(ssp4_highBud_42) = "ssp4_highBud_feat42"
names(ssp4_highBud_51) = "ssp4_highBud_feat51"

# Export
ssp4_highBud_15_rast <- rast(ssp4_highBud_15)
crs(ssp4_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_highBud_15_rast, "data/outputs/shiny/ssp4_highBud_feat15.tif", overwrite = T)

ssp4_highBud_24_rast <- rast(ssp4_highBud_24)
crs(ssp4_highBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_highBud_24_rast, "data/outputs/shiny/ssp4_highBud_feat24.tif", overwrite = T)

ssp4_highBud_33_rast <- rast(ssp4_highBud_33)
crs(ssp4_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_highBud_33_rast, "data/outputs/shiny/ssp4_highBud_feat33.tif", overwrite = T)

ssp4_highBud_42_rast <- rast(ssp4_highBud_42)
crs(ssp4_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_highBud_42_rast, "data/outputs/shiny/ssp4_highBud_feat42.tif", overwrite = T)

ssp4_highBud_51_rast <- rast(ssp4_highBud_51)
crs(ssp4_highBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp4_highBud_51_rast, "data/outputs/shiny/ssp4_highBud_feat51.tif", overwrite = T)


## SSP5 ------------------------------------------------------------------------

## Low budget 
ssp5_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p5)
ssp5_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p5)
ssp5_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p5)
ssp5_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p5)
ssp5_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p5)

# rename rasters
names(ssp5_lowBud_15) = "ssp5_lowBud_feat_15"
names(ssp5_lowBud_24) = "ssp5_lowBud_feat_24"
names(ssp5_lowBud_33) = "ssp5_lowBud_feat_33"
names(ssp5_lowBud_42) = "ssp5_lowBud_feat_42"
names(ssp5_lowBud_51) = "ssp5_lowBud_feat_51"

# Export
ssp5_lowBud_15_rast <- rast(ssp5_lowBud_15)
crs(ssp5_lowBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_lowBud_15_rast, "data/outputs/shiny/ssp5_lowBud_feat15.tif", overwrite = T)

ssp5_lowBud_24_rast <- rast(ssp5_lowBud_24)
crs(ssp5_lowBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_lowBud_24_rast, "data/outputs/shiny/ssp5_lowBud_feat24.tif", overwrite = T)

ssp5_lowBud_33_rast <- rast(ssp5_lowBud_33)
crs(ssp5_lowBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_lowBud_33_rast, "data/outputs/shiny/ssp5_lowBud_feat33.tif", overwrite = T)

ssp5_lowBud_42_rast <- rast(ssp5_lowBud_42)
crs(ssp5_lowBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_lowBud_42_rast, "data/outputs/shiny/ssp5_lowBud_feat42.tif", overwrite = T)

ssp5_lowBud_51_rast <- rast(ssp5_lowBud_51)
crs(ssp5_lowBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_lowBud_51_rast, "data/outputs/shiny/ssp5_lowBud_feat51.tif", overwrite = T)




## High budget
ssp5_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p5)
ssp5_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p5)
ssp5_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p5)
ssp5_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p5)
ssp5_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p5)

# rename rasters
names(ssp5_highBud_15) = "ssp5_highBud_feat15"
names(ssp5_highBud_24) = "ssp5_highBud_feat24"
names(ssp5_highBud_33) = "ssp5_highBud_feat33"
names(ssp5_highBud_42) = "ssp5_highBud_feat42"
names(ssp5_highBud_51) = "ssp5_highBud_feat51"

# Export
ssp5_highBud_15_rast <- rast(ssp5_highBud_15)
crs(ssp5_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_highBud_15_rast, "data/outputs/shiny/ssp5_highBud_feat15.tif", overwrite = T)

ssp5_highBud_24_rast <- rast(ssp5_highBud_24)
crs(ssp5_highBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_highBud_24_rast, "data/outputs/shiny/ssp5_highBud_feat24.tif", overwrite = T)

ssp5_highBud_33_rast <- rast(ssp5_highBud_33)
crs(ssp5_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_highBud_33_rast, "data/outputs/shiny/ssp5_highBud_feat33.tif", overwrite = T)

ssp5_highBud_42_rast <- rast(ssp5_highBud_42)
crs(ssp5_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_highBud_42_rast, "data/outputs/shiny/ssp5_highBud_feat42.tif", overwrite = T)

ssp5_highBud_51_rast <- rast(ssp5_highBud_51)
crs(ssp5_highBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp5_highBud_51_rast, "data/outputs/shiny/ssp5_highBud_feat51.tif", overwrite = T)


## SSP Overlap ----------------------------------------------------------------

## Low budget 
ssp_all_lowBud_15 <- feat_weights(weights = c(1,5), bud = 445.862, p_all)
ssp_all_lowBud_24 <- feat_weights(weights = c(2,4), bud = 445.862, p_all)
ssp_all_lowBud_33 <- feat_weights(weights = c(3,3), bud = 445.862, p_all)
ssp_all_lowBud_42 <- feat_weights(weights = c(4,2), bud = 445.862, p_all)
ssp_all_lowBud_51 <- feat_weights(weights = c(5,1), bud = 445.862, p_all)

# rename rasters
names(ssp_all_lowBud_15) = "ssp_all_lowBud_feat15"
names(ssp_all_lowBud_24) = "ssp_all_lowBud_feat24"
names(ssp_all_lowBud_33) = "ssp_all_lowBud_feat33"
names(ssp_all_lowBud_42) = "ssp_all_lowBud_feat42"
names(ssp_all_lowBud_51) = "ssp_all_lowBud_feat51"

# Export
ssp_all_lowBud_15_rast <- rast(ssp_all_lowBud_15)
crs(ssp_all_lowBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_lowBud_15_rast,
            "data/outputs/shiny/ssp_all_lowBud_feat15.tif", overwrite = T)

ssp_all_lowBud_24_rast <- rast(ssp_all_lowBud_24)
crs(ssp_all_lowBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_lowBud_24_rast, 
            "data/outputs/shiny/ssp_all_lowBud_feat24.tif", overwrite = T)

ssp_all_lowBud_33_rast <- rast(ssp_all_lowBud_33)
crs(ssp_all_lowBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_lowBud_33_rast,
            "data/outputs/shiny/ssp_all_lowBud_feat33.tif", overwrite = T)

ssp_all_lowBud_42_rast <- rast(ssp_all_lowBud_42)
crs(ssp_all_lowBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_lowBud_42_rast, 
            "data/outputs/shiny/ssp_all_lowBud_feat42.tif", overwrite = T)

ssp_all_lowBud_51_rast <- rast(ssp_all_lowBud_51)
crs(ssp_all_lowBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_lowBud_51_rast, 
            "data/outputs/shiny/ssp_all_lowBud_feat51.tif", overwrite = T)



## High budget
ssp_all_highBud_15 <- feat_weights(weights = c(1,5), bud = 3400, p_all)
ssp_all_highBud_24 <- feat_weights(weights = c(2,4), bud = 3400, p_all)
ssp_all_highBud_33 <- feat_weights(weights = c(3,3), bud = 3400, p_all)
ssp_all_highBud_42 <- feat_weights(weights = c(4,2), bud = 3400, p_all)
ssp_all_highBud_51 <- feat_weights(weights = c(5,1), bud = 3400, p_all)

# rename rasters
names(ssp_all_highBud_15) = "ssp_all_highBud_feat15"
names(ssp_all_highBud_24) = "ssp_all_highBud_feat24"
names(ssp_all_highBud_33) = "ssp_all_highBud_feat33"
names(ssp_all_highBud_42) = "ssp_all_highBud_feat42"
names(ssp_all_highBud_51) = "ssp_all_highBud_feat51"

# Export
ssp_all_highBud_15_rast <- rast(ssp_all_highBud_15)
crs(ssp_all_highBud_15_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_highBud_15_rast,
            "data/outputs/shiny/ssp_all_highBud_feat15.tif", overwrite = T)

ssp_all_highBud_24_rast <- rast(ssp_all_highBud_24)
crs(ssp_all_highBud_24_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_highBud_24_rast, 
            "data/outputs/shiny/ssp_all_highBud_feat24.tif", overwrite = T)

ssp_all_highBud_33_rast <- rast(ssp_all_highBud_33)
crs(ssp_all_highBud_33_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_highBud_33_rast, 
            "data/outputs/shiny/ssp_all_highBud_feat33.tif", overwrite = T)

ssp_all_highBud_42_rast <- rast(ssp_all_highBud_42)
crs(ssp_all_highBud_42_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_highBud_42_rast,
            "data/outputs/shiny/ssp_all_highBud_feat42.tif", overwrite = T)

ssp_all_highBud_51_rast <- rast(ssp_all_highBud_51)
crs(ssp_all_highBud_51_rast) <- crs(ssp1_abandoned_crop)
writeRaster(ssp_all_highBud_51_rast,
            "data/outputs/shiny/ssp_all_highBud_feat51.tif", overwrite = T)
```









## FURTHER ANALYSES

Below are misc analyses testing different problem formulations and zonal statistics of their outputs. 

```{r}
# reclass the planning unit layers so each cell has a value of 1
# easier for zonal stats
df <- c(0, 3200, 1)
m <- matrix(df, ncol = 3, byrow = TRUE)

pu1 <- classify(ssp1_pu, m) 
names(pu1) <- "pu1"
pu2 <- classify(ssp2_pu, m)
names(pu2) <- "pu2"
pu3 <- classify(ssp3_pu, m)
names(pu3) <- "pu3"
pu4 <- classify(ssp4_pu, m)
names(pu4) <- "pu4"
pu5 <- classify(ssp5_pu, m)
names(pu5) <- "pu5"
pu_all <- classify(ssp_all_pu, m)
names(pu_all) <- 'pu_all'

# Counts ----------------------------------------------------------------------
## How many pus were available?
ab_pu_ssp1 <- sum(freq(pu1)[3])
ab_pu_ssp2 <- sum(freq(pu2)[3])
ab_pu_ssp3 <- sum(freq(pu3)[3])
ab_pu_ssp4 <- sum(freq(pu4)[3])
ab_pu_ssp5 <- sum(freq(pu5)[3])


## how many pus were selected in each? 
picked_1 <- (freq(ssp1_solution)[2, 3])
picked_2 <- (freq(ssp2_solution)[2, 3])
picked_3 <- (freq(ssp3_solution)[2, 3])
picked_4 <- (freq(ssp4_solution)[2, 3])
picked_5 <- (freq(ssp5_solution)[2, 3])

## percent pus chosen
perc_picked_1 <- (picked_1/(ab_pu_ssp1)) * 100
perc_picked_2 <- (picked_2/(ab_pu_ssp2)) * 100
perc_picked_3 <- (picked_3/(ab_pu_ssp3)) * 100
perc_picked_4 <- (picked_4/(ab_pu_ssp4)) * 100
perc_picked_5 <- (picked_5/(ab_pu_ssp5)) * 100

## how many pus overlapped in all five solutions? 
total_pu <- sum(pu1, pu2, pu3, pu4, pu5)

## how many pus were chosen in all five solutions?
total_picked <- freq(total_solution)[6,3]

```



```{r}
# Zonal Stats------------------------------------------------------------------

### make raster with biome codes as values
biomes_rast <- rasterize(x = biomes_vect,
                         y = ssp1_abandoned_crop,
                         field = "name_biome")       
### remove pantanal biome
biomes_rast_noPant <- mask(biomes_rast, pantanal, inverse = TRUE, updatevalue = NA)

### find how many pus were present in each biome
ssp1_zs_pu_biome <- zonal(pu1, biomes_rast, "sum", na.rm = TRUE)
ssp2_zs_pu_biome <- zonal(pu2, biomes_rast, "sum", na.rm = TRUE)
ssp3_zs_pu_biome <- zonal(pu3, biomes_rast, "sum", na.rm = TRUE)
ssp4_zs_pu_biome <- zonal(pu4, biomes_rast, "sum", na.rm = TRUE)
ssp5_zs_pu_biome <- zonal(pu5, biomes_rast, "sum", na.rm = TRUE)
ssp_all_zs_pu_biome <- zonal(pu_all, biomes_rast, 'sum', na.rm = T)

### combine all the counts into one table
counts_pu_biome <- full_join(ssp1_zs_pu_biome, ssp2_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp3_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp4_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp5_zs_pu_biome, by = "name_biome")



### find how many pus were *prioritized* in each biome
ssp1_zs_sol_biome <- zonal(ssp1_sol, biomes_rast, "sum", na.rm = TRUE)
ssp2_zs_sol_biome <- zonal(ssp2_sol, biomes_rast, "sum", na.rm = TRUE)
ssp3_zs_sol_biome <- zonal(ssp3_sol, biomes_rast, "sum", na.rm = TRUE)
ssp4_zs_sol_biome <- zonal(ssp4_sol, biomes_rast, "sum", na.rm = TRUE)
ssp5_zs_sol_biome <- zonal(ssp5_sol, biomes_rast, "sum", na.rm = TRUE)
ssp_all_zs_sol_biome <- zonal(ssp_all_sol, biomes_rast, "sum", na.rm = TRUE)


#### combine all the counts into one table
counts_sol_biome <- full_join(ssp1_zs_sol_biome, ssp2_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp3_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp4_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp5_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp_all_zs_sol_biome, by = 'name_biome')


## find how many features are in each biome
# convert zero values to nas to mask properly
ssp_all_sol_high[ssp_all_sol_high == 0] <- NA

ssp_all_bio <- mask(bio_ext_brazil, ssp_all_abandoned_crop)
ssp_all_carbon <- mask(carbon_brazil, ssp_all_abandoned_crop)

ssp_all_bio_sol <- mask(ssp_all_bio, ssp_all_sol_high, updatevalue = NA)
ssp_all_carbon_sol <- mask(ssp_all_carbon, ssp_all_sol_high, updatevalue = NA)

ssp_all_zs_bio_biome <- zonal(ssp_all_bio_sol, biomes_rast, 'sum', na.rm = TRUE)










### combine all the counts into one table
counts_sol_state <- full_join(ssp1_zs_sol_biome, ssp2_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp3_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp4_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp5_zs_sol_biome, by = "name_biome")


### find overlap between each ssp
total_pu[total_pu == 5] <- 1
overlap_zs_pu_biome <- zonal(total_pu, biomes_rast, "sum", na.rm = TRUE)

overlap_zs_sol_biome <- zonal(total_solution_overlap, biomes_rast,"sum", na.rm = TRUE)






## By State --------------------------------
states <- read_state(year = 2020, simplified = FALSE)

### create vector for states
states_vect <- states %>%  
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()  

### create raster for states
states_rast <- rasterize(x = states_vect,
                         y = ssp1_abandoned_crop,
                         field = "name_state")  %>% 
  mask(mask = pantanal, inverse = TRUE, updatevalue = NA)

### find how many pus were present in each state
ssp1_zs_pu_state <- zonal(pu1, states_rast, "sum", na.rm = TRUE)
ssp2_zs_pu_state <- zonal(pu2, states_rast, "sum", na.rm = TRUE)
ssp3_zs_pu_state <- zonal(pu3, states_rast, "sum", na.rm = TRUE)
ssp4_zs_pu_state <- zonal(pu4, states_rast, "sum", na.rm = TRUE)
ssp5_zs_pu_state <- zonal(pu5, states_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_pu_state <- full_join(ssp1_zs_pu_state, ssp2_zs_pu_state, by = "name_state") %>% 
  full_join(ssp3_zs_pu_state, by = "name_state") %>% 
  full_join(ssp4_zs_pu_state, by = "name_state") %>% 
  full_join(ssp5_zs_pu_state, by = "name_state")

### find how many pus were *prioritized* in each biome
ssp1_zs_sol_state <- zonal(ssp1_solution, states_rast, "sum", na.rm = TRUE)
ssp2_zs_sol_state <- zonal(ssp2_solution, states_rast, "sum", na.rm = TRUE)
ssp3_zs_sol_state <- zonal(ssp3_solution, states_rast, "sum", na.rm = TRUE)
ssp4_zs_sol_state <- zonal(ssp4_solution, states_rast, "sum", na.rm = TRUE)
ssp5_zs_sol_state <- zonal(ssp5_solution, states_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_sol_state <- full_join(ssp1_zs_sol_state, ssp2_zs_sol_state, by = "name_state") %>% 
  full_join(ssp3_zs_sol_state, by = "name_state") %>% 
  full_join(ssp4_zs_sol_state, by = "name_state") %>% 
  full_join(ssp5_zs_sol_state, by = "name_state")
```


What if budget wasn't a factor? Let's see how much biodiversity and carbon features could be met if 12 million ha (Brazil's goal) of land was restored. To do this, we will convert the costs simply to area and feed these planning units into the prioritization. 

```{r}
#All SSPs with no budget?
noBud_df <- c(0,5, 1)
noBud_m <- matrix(noBud_df, ncol = 3, byrow = TRUE)

ssp_all_pu_noBud <- classify(ssp_all_abandoned_crop, noBud_m)
ssp_all_pu_noBud <- raster(ssp_all_pu_noBud)

p_all_noBud <- problem(ssp_all_pu_noBud, ssp_all_feat) %>% 
  add_relative_targets(c(1, 1)) %>% 
  add_min_shortfall_objective(budget = 120000) %>% 
  add_gurobi_solver(gap = 0.05)

s_all_noBud <- solve(p_all_noBud)

# how many PUs were chosen? answer should be 120,000
eval_n_summary(p_all_noBud, s_all_noBud)
# how much of the targets were met?
eval_target_coverage_summary(p_all_noBud, s_all_noBud)
```


```{r}

### TEST CODE
ssp1_biome_z <- terra::zonal(ssp1_sol$highBud, biomes_rast, 'sum', na.rm = T) %>% 
  rename(pus = 2)
 
reclass_m = matrix(c(0,1,1), ncol = 3, byrow = TRUE)
    X_rcl = terra::classify(ssp1_sol$highBud, reclass_m, include.lowest = TRUE)

ssp1_biome_avail <- terra::zonal(X_rcl, biomes_rast, 'sum', na.rm = T) %>% 
  rename(sol = highBud)

stats_df <- full_join(ssp1_biome_avail, ssp1_biome_z, by = 'name_biome') %>% 
  pivot_longer(cols = 2:3,
               names_to = "category",
               values_to = "amount") %>% 
  mutate(category = factor(category, 
                             levels = c('sol', 'pus')))

ggplot(data = stats_df, aes(x = name_biome, y = amount)) +
  geom_col(aes(fill = category)) +
  theme_minimal()
```


Info on carbon and biodiversity layers
```{r}
# carbon --------------------------------------------------------------------

## mean total carbon by biome
carbon_mean <- zonal(ssp_all_feat, biomes_rast, "mean", na.rm = TRUE)

## how much carbon can be stored in each biome?
carbon_sum <- zonal(carbon_brazil, biomes_rast, 'sum', na.rm = TRUE)

## relative percentage of stored carbon in each biome
carbon_amazon <- (carbon_sum[1,2])/(sum(carbon_sum[2]))
carbon_caatinga <- (carbon_sum[2,2])/(sum(carbon_sum[2]))
carbon_cerrado <- (carbon_sum[3,2])/(sum(carbon_sum[2]))
carbon_atl_for <- (carbon_sum[4,2])/(sum(carbon_sum[2]))
carbon_pampa <- (carbon_sum[5,2])/(sum(carbon_sum[2]))

## how much carbon within the planning units across Brazil?


# Biodiversity --------------------------------------------------------------
## how many pixels in each biome
bio_sum <- zonal(biodiv_binary, biomes_rast, 'sum', na.rm = TRUE)

## relative percentage of bio habitat in each biome
bio_amazon <- (bio_sum[1,2])/(sum(bio_sum[2]))
bio_caatinga <- (bio_sum[2,2])/(sum(bio_sum[2]))
bio_cerrado <- (bio_sum[3,2])/(sum(bio_sum[2]))
bio_atl_for <- (bio_sum[4,2])/(sum(bio_sum[2]))
bio_pampa <- (bio_sum[5,2])/(sum(bio_sum[2]))
```


