---
title: "Prioritization of cropland restoration"
authors: "Nickolas McManus and Michelle Geldin"
date: '2022-10-13'
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(prioritizr)    # prioritization package 
library(gurobi)        # optimizaiton software
library(sf)            # associated with prioritizr
library(terra)         # for reading in and altering rasters
library(raster)        # for formating rasters in prioritizr-friendly format
library(geobr)         # geometries for Brazilian boundaries
library(ggplot2)       # visualize some outputs
library(tidyverse)     # always!
library(janitor)       # cleaning up dfs
```

## Overview
Discuss overall purpose of this analysis. 



## Read in data{.tabset}
First we will read in all the data necessary for formulating and solving the prioritization problem. Broadly, this includes the planning units (locations of projected cropland abandonment), various features (benefits to biodiversity and carbon), and costs of restoration.


### Cropland Abandonment
Read in areas of projected cropland abandonment for five different SSP scenarios in Brazil. These rasters were generated from code located in the `abandoned-cropland` repository, which provides a detailed methodology. Pixel values range from 1 to 7 indicating the frequency of abandonment between 2020 and 2050; values of 1 indicate abandonment only occurred for a five year duration, while pixels with values of 7 were projected to be abandoned for all of 2020-2050.  

Projections of land-use under different SSP scenarios were sourced from:
- Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6

```{r}
# read in the cropland abandonment
ssp1_abandoned_crop <- rast("data/inputs/ssp1/SSP1_abandoned_cropland_brazil.tif")
ssp2_abandoned_crop <- rast("data/inputs/ssp2/SSP2_abandoned_cropland_brazil.tif")
ssp3_abandoned_crop <- rast("data/inputs/ssp3/SSP3_abandoned_cropland_brazil.tif")
ssp4_abandoned_crop <- rast("data/inputs/ssp4/SSP4_abandoned_cropland_brazil.tif")
ssp5_abandoned_crop <- rast("data/inputs/ssp5/SSP5_abandoned_cropland_brazil.tif")
```


### Brazilian Biomes
The boundary information for Brazilian biomes is read-in using the geobr package. Biome polygons are sourced from the Brazilian Institute of Geography and Statistics (IBGE). This dataset uses 2019 IBGE data at scale 1:250.000.

Source: https://www.ibge.gov.br/geociencias/cartas-e-mapas/informacoes-ambientais/15842-biomas.html?=&t=acesso-ao-produto

```{r}
# read in biome boundaries from geobr package
biomes <- read_biomes(
  year = 2019,         #the most recent data available
  simplified = FALSE   #full resolution for boundaries
)

# create vector for whole of Brazil
biomes_vect <- biomes[-7, ] %>%  #remove coastal system biome
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% #change crs to match other rasters
  vect()  #turn sf into vector; easier to crop rasts with


# create vectors for each of the biomes -----------------------------------------
amazon <- biomes %>% 
  filter(code_biome == 1) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

caatinga <- biomes %>% 
  filter(code_biome == 2) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

cerrado <- biomes %>% 
  filter(code_biome == 3) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

atlantic_forest <- biomes %>% 
  filter(code_biome == 4) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pampa <- biomes %>% 
  filter(code_biome == 5) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pantanal <- biomes %>% 
  filter(code_biome == 6) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()
```


### Carbon
This dataset, originally created by Cook-Patton et al. and updated by Global Forest Watch, estimates the rate of carbon sequestration in aboveground and belowground live biomass during the first 30 years of natural forest regeneration. Spatial sequestration estimates include all forest and savanna biomes in units of MgC/ha/yr at a ~1 km resolution. This carbon data is cropped, masked, and reprojected to match the spatial extent and resolution of other rasters used in the prioritization. Because sequestration data is mostly missing for the Pantanal biome in Brazil, it will be removed from this and all data in the analysis. 

Source: https://data.globalforestwatch.org/documents/gfw::carbon-accumulation-potential-from-natural-forest-regrowth-in-forest-and-savanna-biomes/about

```{r, eval=FALSE}
#NOTE: this code chunk defaults to not evaluate.
#only needs to be run once and cropped raster is saved

# read in carbon layer
carbon_global <- rast("data/inputs/sequestration_rate_mean_aboveground_full_extent_Mg_C_ha_yr.tif")

#cannot crop/mask carbon data when biomes are different crs
#change biome crs because it's faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(carbon_global))

# crop carbon data to Brazil
carbon_brazil <- carbon_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>%  
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect)) 

# default resolution of carbon layer is 898.8m
# change resolution to match other rasters (1km)
carbon_brazil <- resample(carbon_brazil, ssp1_abandoned_crop, method = "bilinear")

# remove Panatnal biome from carbon layer
carbon_brazil <- mask(carbon_brazil, pantanal, inverse = TRUE, updatevalue = NA, touches = TRUE)

# change the name saved within raster for easier recall
# after stacking feature rasters
names(carbon_brazil) <- "carbon_brazil"

# test plot to visualize results
plot(carbon_brazil)
plot(biomes_vect, add = TRUE)

#export carbon_brazil as intermediate raster
writeRaster(carbon_brazil, "data/outputs/carbon_seq_rate_Brazil.tif", overwrite = TRUE)
```

```{r}
# run below line of code just to read in layer if already created
carbon_brazil <- rast("data/outputs/carbon_seq_rate_Brazil.tif")
```


### Biodiversity
Read in biodiversity data. Info about the data and where it is from. 

After reading in the global layer, we will crop it to Brazil, resample to match the resolution of the planning units, and save the output raster. 

```{r, eval=FALSE}
#NOTE: this code chunk defaults to not evaluate.
#only needs to be run once and then cropped raster is saved

#read in global biodiversity data
biodiversity_global <- rast("data/inputs/sparc_conservationPriorities.tif")


#temporarily change biome crs to crop/mask biodiversity data
#faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(biodiversity_global))


# crop biodiversity layer to Brazil
biodiversity_brazil <- biodiversity_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect))


# change resolution of biodiversity layer (4.72km) to match other rasters (1km)
# first use terra:disagg()
biodiversity_brazil <- disagg(biodiversity_brazil, 
                              fact = 5,              
                              method = "bilinear")  
# resample higher res biodiversity to ensure same res and geometries 
# (fine-tuning the res)
biodiversity_brazil <- resample(biodiversity_brazil, ssp1_abandoned_crop, method = "bilinear")

# remove Pantanal biome from biodiversity layer
biodiversity_brazil <- mask(biodiversity_brazil, pantanal, inverse = TRUE, updatevalue = NA)

# change the name saved within raster for easier recall
# after stacking feature rasters
names(biodiversity_brazil) <- "biodiversity_brazil"

# test plot to visualize results
plot(biodiversity_brazil)
plot(biomes_vect, add = TRUE)

#export biodiversity_brazil as intermediate raster
writeRaster(biodiversity_brazil, "data/outputs/sparc_conservationPriorities_Brazil.tif", overwrite = TRUE)
```

```{r}
# run below line of code just to read in layer if already created
biodiversity_brazil <- rast("data/outputs/sparc_conservationPriorities_Brazil.tif")
```


### Cost Layer
The cost of restoration in Brazil differs with location (by biome), restoration method (six types), and environmental condition of land (favorable or unfavorable). These costs were sourced from a 2017 Nature Conservancy report, written by Rubens de Miranda Benini and Sergio Adeodato. 

```{r}
#read in cost data
costs <- read_csv("data/inputs/tnc_cost_biomes_methods.csv") %>% 
  clean_names()
```












## Prep data for prioritization

Wrangle data into formats suitable for prioritizr inputs. 

### Cost
Select the restoration method and environmental condition you want to investigate for the cost layer. Enter these in the `filter` lines of code. 

- NOTE: need to multiple cost by 100 and rerun all code to reflect cost per km rather than hectare
```{r}
# Looking at conducting natural regeneration and "unfavorable environmental conditions" (CAD)
costs_filtered <- costs %>% 
  filter(restoration_method == "conducting_natural_regeneration") %>% 
  filter(environmental_conditions == "CAD") %>% 
  filter(biome_code != 6) %>%   #remove pantanal data
  select(!c(id, biome))  #remove id and biome columns
```

To properly assign costs to our planning units, we need to merge our biome data with the filtered cost data. This dataset will then be rasterized for the next step of finalizing our planning units. 
```{r}
# first join biomes with cost layer 
biomes_cost <- biomes %>% 
  full_join(costs_filtered, by = c("code_biome" = "biome_code"))

# turn cost/biome data into a vector
biomes_cost_vect <- biomes_cost[-7, ] %>%   #remove coastal system biome
  st_transform(crs = crs(ssp1_abandoned_crop)) %>%  #change crs to match other rasters
  vect()  

# turn vect into a raster
biomes_cost_rast <- rasterize(x = biomes_cost_vect,      #vect file to rasterize
                              y = ssp1_abandoned_crop,   #raster file to provide example crs and spatial extent
                              field = "cost_r_ha")       #values for raster

# remove pantanal biome
biomes_cost_rast <- mask(biomes_cost_rast, pantanal, inverse = TRUE, updatevalue = NA)
```


### Planning Units
Decide on the restoration method and land status desired to analyze the cost, then merge to assign costs to each planning unit.

```{r}
# mask the cost by biomes to abandoned cropland raster
# this will isolate pixels of interest (planning units) with associated cost value

ssp1_pu <- mask(biomes_cost_rast, ssp1_abandoned_crop)
ssp1_pu <- raster(ssp1_pu)  # turn into prioritizr-friendly format

ssp2_pu <- mask(biomes_cost_rast, ssp2_abandoned_crop)
ssp2_pu <- raster(ssp2_pu)

ssp3_pu <- mask(biomes_cost_rast, ssp3_abandoned_crop)
ssp3_pu <- raster(ssp3_pu)

ssp4_pu <- mask(biomes_cost_rast, ssp4_abandoned_crop)
ssp4_pu <- raster(ssp4_pu)

ssp5_pu <- mask(biomes_cost_rast, ssp5_abandoned_crop)
ssp5_pu <- raster(ssp5_pu)
```




### Features

We want to break up biodiversity layer into quantiles, then keep only the top by making it binary. 
Then we can set targets for meeting a percentage of the "available" biodiversity area
```{r}
##biodiversity_brazil_quant <- global(biodiversity_brazil, probs=seq(0, 1, 0.25), na.rm = TRUE)

#biodiv_binary <- rast("data/inputs/BinaryBiodiversity.tif")

#biodiv_binary <- mask(biodiv_binary, biomes_vect)
#biodiv_binary <- mask(biodiv_binary, pantanal, inverse = TRUE, updatevalue = NA)

#writeRaster(biodiv_binary, "data/outputs/biodiversity_binary_brazil.tif")

biodiv_binary <- rast("data/outputs/biodiversity_binary_brazil.tif")

#plot(biodiv_binary)
  
```
 

To evaluate carbon and biodiversity as features in the prioritization problem, they must be merged into a raster stack.
```{r}
# convert biodiversity and carbon layers from
# terra "SpatRaster" format to raster "RasterLayer"
biodiv_binary <- raster(biodiv_binary)
carbon_brazil <- raster(carbon_brazil)

# now stack the two rasters
features <- stack(biodiv_binary, carbon_brazil)

# set up targets
## 60,000km for biodiversity and 20,000Mg carbon
targets <- c(50000, 450000)
```





## Problem Formulations

This is where we will start formulating the problem to be solved in prioritizr!
```{r}
# SSP1 --------------------------------------------
p1 <- problem(ssp1_pu, features) %>% 
  # minimize cost while ensuring all targets are met
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP2 --------------------------------------------
p2 <- problem(ssp2_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_gurobi_solver(gap = 0.05) 


# SSP3 --------------------------------------------
p3 <- problem(ssp3_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP4 --------------------------------------------
p4 <- problem(ssp4_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP5 --------------------------------------------
p5 <- problem(ssp5_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_gurobi_solver(gap = 0.05) 

```
 


## Problem Solving
```{r}
s1 <- solve(p1)

s2 <- solve(p2)

s3 <- solve(p3)

s4 <- solve(p4)

s5 <- solve(p5)

## convert RasterLayers back to SpatRast and export
ssp1_solution <- rast(s1)
crs(ssp1_solution) = crs(ssp1_abandoned_crop)
names(ssp1_solution) <- "ssp1_solution"
writeRaster(ssp1_solution, "data/outputs/ssp1_solution.tif", overwrite = TRUE)

ssp2_solution <- rast(s2)
crs(ssp2_solution) = crs(ssp1_abandoned_crop)
names(ssp2_solution) <- "ssp2_solution"
writeRaster(ssp2_solution, "data/outputs/ssp2_solution.tif", overwrite = TRUE)

ssp3_solution <- rast(s3)
crs(ssp3_solution) = crs(ssp1_abandoned_crop)
names(ssp3_solution) <- "ssp3_solution"
writeRaster(ssp3_solution, "data/outputs/ssp3_solution.tif", overwrite = TRUE)

ssp4_solution <- rast(s4)
crs(ssp4_solution) = crs(ssp1_abandoned_crop)
names(ssp4_solution) <- "ssp4_solution"
writeRaster(ssp4_solution, "data/outputs/ssp4_solution.tif", overwrite = TRUE)

ssp5_solution <- rast(s5)
crs(ssp5_solution) = crs(ssp1_abandoned_crop)
names(ssp5_solution) <- "ssp5_solution"
writeRaster(ssp5_solution, "data/outputs/ssp5_solution.tif", overwrite = TRUE)

total_solution <- ssp1_solution + ssp2_solution + ssp3_solution + ssp4_solution + ssp5_solution
writeRaster(total_solution, "data/outputs/total_solution.tif", overwrite = TRUE)
```


Experimenting with boundary length modifiers:
What if we do it with BLM of 50?
```{r, eval= FALSE}
# SSP1 --------------------------------------------
p1_blm50 <- problem(ssp1_pu, features) %>% 
  # minimize cost while ensuring all targets are met
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_boundary_penalties(penalty = 50) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP2 --------------------------------------------
p2_blm50 <- problem(ssp2_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_boundary_penalties(penalty = 50) %>% 
  add_gurobi_solver(gap = 0.05) 


# SSP3 --------------------------------------------
p3_blm50 <- problem(ssp3_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_boundary_penalties(penalty = 50) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP4 --------------------------------------------
p4_blm50 <- problem(ssp4_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_boundary_penalties(penalty = 50) %>% 
  add_gurobi_solver(gap = 0.05)


# SSP5 --------------------------------------------
p5_blm50 <- problem(ssp5_pu, features) %>% 
  add_min_set_objective() %>% 
  add_absolute_targets(targets) %>% 
  add_boundary_penalties(penalty = 50) %>% 
  add_gurobi_solver(gap = 0.05) 


s1_blm50 <- solve(p1)

s2_blm50 <- solve(p2)

s3_blm50 <- solve(p3)

s4_blm50 <- solve(p4)

s5_blm50 <- solve(p5)

```


## Further analyses

Here we'll compare the prioritizr solutions with further analyses. 
Things to-do:
- create nice table of results
- update costs

```{r}
# reclass the planning unit layers so each cell has a value of 1
# easier for zonal stats
df <- c(1, 3200, 1)
m <- matrix(df, ncol = 3, byrow = TRUE)

pu1 <- classify(ssp1_pu, m) 
names(pu1) <- "pu1"
pu2 <- classify(ssp2_pu, m)
names(pu2) <- "pu2"
pu3 <- classify(ssp3_pu, m)
names(pu3) <- "pu3"
pu4 <- classify(ssp4_pu, m)
names(pu4) <- "pu4"
pu5 <- classify(ssp5_pu, m)
names(pu5) <- "pu5"

# Counts ----------------------------------------------------------------------
## How many pus were available?
ab_pu_ssp1 <- sum(freq(pu1)[3])
ab_pu_ssp2 <- sum(freq(pu2)[3])
ab_pu_ssp3 <- sum(freq(pu3)[3])
ab_pu_ssp4 <- sum(freq(pu4)[3])
ab_pu_ssp5 <- sum(freq(pu5)[3])


## how many pus were selected in each? 
picked_1 <- (freq(ssp1_solution)[2, 3])
picked_2 <- (freq(ssp2_solution)[2, 3])
picked_3 <- (freq(ssp3_solution)[2, 3])
picked_4 <- (freq(ssp4_solution)[2, 3])
picked_5 <- (freq(ssp5_solution)[2, 3])

## percent pus chosen
perc_picked_1 <- (picked_1/(ab_pu_ssp1)) * 100
perc_picked_2 <- (picked_2/(ab_pu_ssp2)) * 100
perc_picked_3 <- (picked_3/(ab_pu_ssp3)) * 100
perc_picked_4 <- (picked_4/(ab_pu_ssp4)) * 100
perc_picked_5 <- (picked_5/(ab_pu_ssp5)) * 100

## how many pus overlapped in all five solutions? 
total_pu <- sum(pu1, pu2, pu3, pu4, pu5)

## how many pus were chosen in all five solutions?
total_picked <- freq(total_solution)[6,3]



# Costs ----------------------------------------------------------------------
## NOTE: these all need to be multiplied by 100 to get cost/km2
## return cost for each ssp 
ssp1_cost <- eval_cost_summary(p1, s1)
ssp2_cost <- eval_cost_summary(p2, s2)
ssp3_cost <- eval_cost_summary(p3, s3)
ssp4_cost <- eval_cost_summary(p4, s4)
ssp5_cost <- eval_cost_summary(p5, s5)

## return cost of total solution
df2 <- c(0, 4, NA,
         4, 6, 1)
m2<- matrix(df2, ncol =3, byrow = TRUE)
total_solution_overlap <- classify(total_solution, m2, include.lowest  = TRUE)

total_cost <- mask(biomes_cost_rast, total_solution_overlap)
total_cost_int <- (freq(total_cost)[1,2] * freq(total_cost)[1,3]) + 
  (freq(total_cost)[2,2] * freq(total_cost)[2,3]) + 
  (freq(total_cost)[3,2] * freq(total_cost)[3,3]) +
  (freq(total_cost)[4,2] * freq(total_cost)[4,3]) +
  (freq(total_cost)[5,2] * freq(total_cost)[5,3])


# Zonal Stats------------------------------------------------------------------

## By Biome --------------------------------
### make raster with biome codes as values
biomes_rast <- rasterize(x = biomes_cost_vect,
                         y = ssp1_abandoned_crop,
                         field = "name_biome")       
### remove pantanal biome
biomes_rast <- mask(biomes_rast, pantanal, inverse = TRUE, updatevalue = NA)

### find how many pus were present in each biome
ssp1_zs_pu_biome <- zonal(pu1, biomes_rast, "sum", na.rm = TRUE)
ssp2_zs_pu_biome <- zonal(pu2, biomes_rast, "sum", na.rm = TRUE)
ssp3_zs_pu_biome <- zonal(pu3, biomes_rast, "sum", na.rm = TRUE)
ssp4_zs_pu_biome <- zonal(pu4, biomes_rast, "sum", na.rm = TRUE)
ssp5_zs_pu_biome <- zonal(pu5, biomes_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_pu_biome <- full_join(ssp1_zs_pu_biome, ssp2_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp3_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp4_zs_pu_biome, by = "name_biome") %>% 
  full_join(ssp5_zs_pu_biome, by = "name_biome")

### find how many pus were *prioritized* in each biome
ssp1_zs_sol_biome <- zonal(ssp1_solution, biomes_rast, "sum", na.rm = TRUE)
ssp2_zs_sol_biome <- zonal(ssp2_solution, biomes_rast, "sum", na.rm = TRUE)
ssp3_zs_sol_biome <- zonal(ssp3_solution, biomes_rast, "sum", na.rm = TRUE)
ssp4_zs_sol_biome <- zonal(ssp4_solution, biomes_rast, "sum", na.rm = TRUE)
ssp5_zs_sol_biome <- zonal(ssp5_solution, biomes_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_sol_state <- full_join(ssp1_zs_sol_biome, ssp2_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp3_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp4_zs_sol_biome, by = "name_biome") %>% 
  full_join(ssp5_zs_sol_biome, by = "name_biome")


### find overlap between each ssp
total_pu[total_pu == 5] <- 1
overlap_zs_pu_biome <- zonal(total_pu, biomes_rast, "sum", na.rm = TRUE)

overlap_zs_sol_biome <- zonal(total_solution_overlap, biomes_rast,"sum", na.rm = TRUE)

## By State --------------------------------
states <- read_state(year = 2020, simplified = FALSE)

### create vector for states
states_vect <- states %>%  
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()  

### create raster for states
states_rast <- rasterize(x = states_vect,
                         y = ssp1_abandoned_crop,
                         field = "name_state")  %>% 
  mask(mask = pantanal, inverse = TRUE, updatevalue = NA)

### find how many pus were present in each state
ssp1_zs_pu_state <- zonal(pu1, states_rast, "sum", na.rm = TRUE)
ssp2_zs_pu_state <- zonal(pu2, states_rast, "sum", na.rm = TRUE)
ssp3_zs_pu_state <- zonal(pu3, states_rast, "sum", na.rm = TRUE)
ssp4_zs_pu_state <- zonal(pu4, states_rast, "sum", na.rm = TRUE)
ssp5_zs_pu_state <- zonal(pu5, states_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_pu_state <- full_join(ssp1_zs_pu_state, ssp2_zs_pu_state, by = "name_state") %>% 
  full_join(ssp3_zs_pu_state, by = "name_state") %>% 
  full_join(ssp4_zs_pu_state, by = "name_state") %>% 
  full_join(ssp5_zs_pu_state, by = "name_state")

### find how many pus were *prioritized* in each biome
ssp1_zs_sol_state <- zonal(ssp1_solution, states_rast, "sum", na.rm = TRUE)
ssp2_zs_sol_state <- zonal(ssp2_solution, states_rast, "sum", na.rm = TRUE)
ssp3_zs_sol_state <- zonal(ssp3_solution, states_rast, "sum", na.rm = TRUE)
ssp4_zs_sol_state <- zonal(ssp4_solution, states_rast, "sum", na.rm = TRUE)
ssp5_zs_sol_state <- zonal(ssp5_solution, states_rast, "sum", na.rm = TRUE)

### combine all the counts into one table
counts_sol_state <- full_join(ssp1_zs_sol_state, ssp2_zs_sol_state, by = "name_state") %>% 
  full_join(ssp3_zs_sol_state, by = "name_state") %>% 
  full_join(ssp4_zs_sol_state, by = "name_state") %>% 
  full_join(ssp5_zs_sol_state, by = "name_state")
```

