---
title: "Prioritization of cropland restoration"
authors: "Nickolas McManus and Michelle Geldin"
date: '2022-10-13'
output: 
 html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
    code_folding: hide
    smooth_scroll: yes
    collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(prioritizr)    # optimization software! 
library(terra)         # for reading in and altering rasters
library(sf)            
library(raster)        # for formating rasters in prioritizr-friendly format
library(geobr)         # geometries for Brazilian biomes
library(ggplot2)       # visualize some outputs
library(tidyverse)     # always!
library(janitor)
```

## Overview
Discuss overall purpose of this analysis. 



## Read in data{.tabset}
First we will read in all the data necessary for formulating and solving the prioritization problem. Broadly, this includes the planning units (locations of projected cropland abandonment), various features (benefits to biodiversity and carbon), and costs of restoration.


### Cropland Abandonment (Rasters)
Read in areas of projected cropland abandonment for five different SSP scenarios in Brazil. These rasters were generated from code located in the `abandoned-cropland` repository, which provides a detailed methodology. Pixel values range from 1 to 7 indicating the frequency of abandonment between 2020 and 2050; values of 1 indicate abandonment only occurred for a five year duration, while pixels with values of 7 were projected to be abandoned for all of 2020-2050.  

Projections of land-use under different SSP scenarios were sourced from:
- Chen, G., Li, X., & Liu, X. (2022). Global land projection based on plant functional types with a 1-km resolution under socio-climatic scenarios. Scientific Data, 9(1), 125. https://doi.org/10.1038/s41597-022-01208-6

```{r}
# read in the cropland abandonment
ssp1_abandoned_crop <- rast("data/inputs/ssp1/SSP1_abandoned_cropland_brazil.tif")
ssp2_abandoned_crop <- rast("data/inputs/ssp2/SSP2_abandoned_cropland_brazil.tif")
ssp3_abandoned_crop <- rast("data/inputs/ssp3/SSP3_abandoned_cropland_brazil.tif")
ssp4_abandoned_crop <- rast("data/inputs/ssp4/SSP4_abandoned_cropland_brazil.tif")
ssp5_abandoned_crop <- rast("data/inputs/ssp5/SSP5_abandoned_cropland_brazil.tif")
```


### Planning Units (Shapefiles)
These shapefiles were produced in ArcGIS Pro by polygonizing the above rasters and then spatially joining with Brazilian biome shapefile. Here we will read in the shapefiles and clean up their dataframe. 

- *This section may or may not get deleted in future if we only need rasters for problem formulation*

```{r, eval=FALSE}
# read in planning units for SSP1
ssp1_pu <- st_read("data/inputs/ssp1/ssp1_pu.shp") %>% 
  clean_names() %>% 
  
  # only keep select columns of interest
  select(join_count, target_fid, bioma, cd_bioma, shape_area) %>% 
  
  # rename some columns to make later merging easier
  rename(id = target_fid,
         biome_code = cd_bioma,
         biome = bioma) %>%
  
  #remove pus that are located outside Brazil biome boundaries
  filter(join_count != 0)

# read in planning units for SSP2 -------------------------------------------
ssp2_pu <- read_sf("data/inputs/ssp2/ssp2_pu.shp") %>% 
  clean_names() %>% 
  select(join_count, target_fid, bioma, cd_bioma, shape_area) %>% 
  rename(id = target_fid,
         biome_code = cd_bioma,
         biome = bioma) %>%
  filter(join_count != 0)

# read in planning units for SSP3 -------------------------------------------
ssp3_pu <- read_sf("data/inputs/ssp3/ssp3_pu.shp") %>% 
  clean_names() %>% 
  select(join_count, target_fid, bioma, cd_bioma, shape_area) %>% 
  rename(id = target_fid,
         biome_code = cd_bioma,
         biome = bioma) %>%
  filter(join_count != 0)

# read in planning units for SSP4 -------------------------------------------
ssp4_pu <- read_sf("data/inputs/ssp4/ssp4_pu.shp") %>% 
  clean_names() %>% 
  select(join_count, target_fid, bioma, cd_bioma, shape_area) %>% 
  rename(id = target_fid,
         biome_code = cd_bioma,
         biome = bioma) %>%
  filter(join_count != 0)

# read in planning units for SSP5 -------------------------------------------
ssp5_pu <- read_sf("data/inputs/ssp5/ssp5_pu.shp") %>% 
  clean_names() %>% 
  select(join_count, target_fid, bioma, cd_bioma, shape_area) %>% 
  rename(id = target_fid,
         biome_code = cd_bioma,
         biome = bioma) %>%
  filter(join_count != 0)
```


### Brazilian Biomes
The boundary information for Brazilian biomes is read-in using the geobr package. Biome polygons are sourced from the Brazilian Institute of Geography and Statistics (IBGE). This dataset uses 2019 IBGE data at scale 1:250.000.

Source: https://www.ibge.gov.br/geociencias/cartas-e-mapas/informacoes-ambientais/15842-biomas.html?=&t=acesso-ao-produto

```{r}
# read in biome boundaries from geobr package
biomes <- read_biomes(
  year = 2019,         #the most recent data available
  simplified = FALSE   #full resolution for boundaries
)

# create vector for whole of Brazil
biomes_vect <- biomes[-7, ] %>%  #remove coastal system biome
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% #change crs to match other rasters
  vect()  #turn sf into vector; easier to crop rasts with


# create vectors for each of the biomes -----------------------------------------
amazon <- biomes %>% 
  filter(code_biome == 1) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

caatinga <- biomes %>% 
  filter(code_biome == 2) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

cerrado <- biomes %>% 
  filter(code_biome == 3) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

atlantic_forest <- biomes %>% 
  filter(code_biome == 4) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pampa <- biomes %>% 
  filter(code_biome == 5) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()

pantanal <- biomes %>% 
  filter(code_biome == 6) %>% 
  st_transform(crs = crs(ssp1_abandoned_crop)) %>% 
  vect()
```


### Carbon
This dataset, originally created by Cook-Patton et al. and updated by Global Forest Watch, estimates the rate of carbon sequestration in aboveground and belowground live biomass during the first 30 years of natural forest regeneration. Spatial sequestration estimates include all forest and savanna biomes in units of MgC/ha/yr at a ~1 km resolution. This carbon data is cropped, masked, and reprojected to match the spatial extent and resolution of other rasters used in the prioritization. 

Source: https://data.globalforestwatch.org/documents/gfw::carbon-accumulation-potential-from-natural-forest-regrowth-in-forest-and-savanna-biomes/about

```{r, eval=FALSE}
#NOTE: this code chunk defaults to not evaluate.
#only needs to be run once and cropped raster is saved

# read in carbon layer
carbon_global <- rast("data/inputs/sequestration_rate__mean__aboveground__full_extent__Mg_C_ha_yr.tif")

#cannot crop/mask carbon data when biomes are different crs
#change biome crs because it's faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(carbon_global))

# crop carbon data to Brazil
carbon_brazil <- carbon_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>%  
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect)) 

# default resolution of carbon layer is 898.8m
# change resolution to match other rasters (1km)
carbon_brazil <- resample(carbon_brazil, ssp1_abandoned_crop, method = "bilinear")

# change the name saved within raster for easier recall
# after stacking feature rasters
names(carbon_brazil) <- "carbon_brazil"

# test plot to visualize results
plot(carbon_brazil)

#export carbon_brazil as intermediate raster
writeRaster(carbon_brazil, "data/outputs/carbon_seq_rate_Brazil.tif")
```

```{r}
# run below line of code just to read in layer if already created
carbon_brazil <- rast("data/outputs/carbon_seq_rate_Brazil.tif")
```


### Biodiversity
Read in biodiversity data. Info about the data and where it is from. 

After reading in the global layer, we will crop it to Brazil, resample to match the resolution of the planning units, and save the output raster. 

```{r, eval=FALSE}
#NOTE: this code chunk defaults to not evaluate.
#only needs to be run once and then cropped raster is saved

#read in global biodiversity data
biodiversity_global <- rast("data/inputs/sparc_conservationPriorities.tif")


#temporarily change biome crs to crop/mask biodiversity data
#faster than changing crs of global raster
brazil_trans <- biomes_vect %>% 
  project(y = crs(biodiversity_global))


# crop biodiversity layer to Brazil
biodiversity_brazil <- biodiversity_global %>% 
  crop(brazil_trans) %>% 
  mask(brazil_trans) %>% 
  #change crs back to match other rasters
  terra::project(y = crs(biomes_vect))


# change resolution of biodiversity layer (4.72km) to match other rasters (1km)
# first use terra:disagg()
biodiversity_brazil <- disagg(biodiversity_brazil, 
                              fact = 5,              
                              method = "bilinear")  
# resample higher res biodiversity to ensure same res and geometries 
# (fine-tuning the res)
biodiversity_brazil <- resample(biodiversity_brazil, ssp1_abandoned_crop, method = "bilinear")


# change the name saved within raster for easier recall
# after stacking feature rasters
names(biodiversity_brazil) <- "biodiversity_brazil"

# test plot to visualize results
plot(biodiversity_brazil)

#export biodiversity_brazil as intermediate raster
writeRaster(biodiversity_brazil, "data/outputs/sparc_conservationPriorities_Brazil.tif")
```

```{r}
# run below line of code just to read in layer if already created
biodiversity_brazil <- rast("data/outputs/sparc_conservationPriorities_Brazil.tif")
```


### Cost Layer
The cost of restoration in Brazil differs with location (by biome), restoration method (six types), and environmental condition of land (favorable or unfavorable). These costs were sourced from a 2017 Nature Conservancy report, written by Rubens de Miranda Benini and Sergio Adeodato. 

```{r}
#read in cost data
costs <- read_csv("data/inputs/tnc_cost_biomes_methods.csv") %>% 
  clean_names()
```


## Prepare data for prioritization

### Cost
```{r}
# select the restoration method and land status you want to use

# Looking at conducting natural regeneration and "unfavorable environmental conditions" (CAD)
costs_filtered <- costs %>% 
  filter(restoration_method == "conducting_natural_regeneration") %>% 
  filter(environmental_conditions == "CAD") %>% 
  select(!c(id, biome))  #remove id and biome columns
```


### Planning Units
Decide on the restoration method and land status desired to analyze the cost, then merge to assign costs to each planning unit.

```{r}
# merging costs with pu rasters by reassigning value of rasters

#Steps:
##1. crop ssp1 layer by individual biomes (turn 1 raster into 6)
##2. reassign values within each biome to costs (can this be automatically pulled from csv??)
##3. merge/mosaic rasters back together into one ssp raster with updated values

#-------------------------------------------------------------------------------
#Test with SSP1:

##Step 1: crop ssp1 layer by each biome to make 6 new rasters
ssp1_amazon <- crop(ssp1_abandoned_crop, amazon, mask = TRUE, touches = TRUE)

ssp1_caatinga <- crop(ssp1_abandoned_crop, caatinga, mask = TRUE, touches = TRUE)

ssp1_cerrado <- crop(ssp1_abandoned_crop, cerrado, mask = TRUE, touches = TRUE)

ssp1_atlantic_forest <- crop(ssp1_abandoned_crop, atlantic_forest, mask = TRUE, touches = TRUE)

ssp1_pampa <- crop(ssp1_abandoned_crop, pampa, mask = TRUE, touches = TRUE)

ssp1_pantanal <-crop(ssp1_abandoned_crop, pantanal, mask = TRUE, touches = TRUE)

##Step 2: unique costs for each biome based on what restoration method was selected
cost_amazon <- costs_filtered$cost_r_ha[1] 
cost_caatinga <- costs_filtered$cost_r_ha[2]
cost_cerrado <- costs_filtered$cost_r_ha[3]
cost_atlantic_forest <- costs_filtered$cost_r_ha[4]
cost_pampa <- costs_filtered$cost_r_ha[5]
cost_pantanal <- costs_filtered$cost_r_ha[6]

# reclassify biomes
amazon_m <- matrix(data = c(1, 7, cost_amazon), ncol = 3, byrow = TRUE)
ssp1_amazon <- classify(ssp1_amazon, amazon_m, include.lowest=TRUE)

caatinga_m <- matrix(data = c(1, 7, cost_caatinga), ncol = 3, byrow = TRUE)
ssp1_caatinga <- classify(ssp1_caatinga, caatinga_m, include.lowest=TRUE)

cerrado_m <- matrix(data = c(1, 7, cost_cerrado), ncol = 3, byrow = TRUE)
ssp1_cerrado <- classify(ssp1_cerrado, cerrado_m, include.lowest=TRUE)

atlantic_forest_m <- matrix(data = c(1, 7, cost_atlantic_forest), ncol = 3, byrow = TRUE)
ssp1_atlantic_forest <- classify(ssp1_atlantic_forest, atlantic_forest_m, include.lowest=TRUE)

pampa_m <- matrix(data = c(1, 7, cost_pampa), ncol = 3, byrow = TRUE)
ssp1_pampa <- classify(ssp1_pampa, pampa_m, include.lowest=TRUE)

pantanal_m <- matrix(data = c(1, 7, cost_pantanal), ncol = 3, byrow = TRUE)
ssp1_pantanal <- classify(ssp1_pantanal, pantanal_m, include.lowest=TRUE)

##Step 3: Merge biome rasters back together
# create SpatRaster collection of reclassified biomes
ssp1_list <- sprc(ssp1_amazon, ssp1_caatinga, ssp1_cerrado, ssp1_atlantic_forest, ssp1_pampa, ssp1_pantanal)

# merge back into one raster for Brazil
ssp1_abandoned_crop_cost <- merge(ssp1_list) 
ssp1_abandoned_crop_cost <- ssp1_abandoned_crop_cost %>%
  crop(biomes_vect) %>% 
  mask(biomes_vect)

# make sure extents match for priotizir problem
ssp1_pu <- resample(ssp1_abandoned_crop_cost, ssp1_abandoned_crop, method = "near")

# turn into prioritizr-friendly format
ssp1_pu <- raster(ssp1_abandoned_crop_cost)

## try changing touches and masking within crop function to TRUE

```





- *section may be removed if shapefiles no longer used*
```{r}
# select the restoration method and land status you want to use

# Looking at conducting natural regeneration and "unfavorable environmental conditions" (CAD)
costs_filtered <- costs %>% 
  filter(restoration_method == "conducting_natural_regeneration") %>% 
  filter(environmental_conditions == "CAD") %>% 
  select(!c(id, biome))  #remove id and biome columns


# now merge the filtered cost with planning units for each SSP

ssp1_pu_costs <- ssp1_pu %>% 
  full_join(costs_cnr_cad, by = "biome_code") %>%  #merge objects
  relocate(geometry, .after = environmental_conditions)    #moves geometry column to end
 #st_drop_geometry()   #may want to just drop geometry altogether? 

ssp2_pu_costs <- ssp2_pu %>% 
  full_join(costs_cnr_cad, by = c("biome_code")) %>% 
  relocate(geometry, .after = environmental_conditions)   

ssp3_pu_costs <- ssp3_pu %>% 
  full_join(costs_cnr_cad, by = c("biome_code")) %>% 
  relocate(geometry, .after = environmental_conditions)  

ssp4_pu_costs <- ssp4_pu %>% 
  full_join(costs_cnr_cad, by = c("biome_code")) %>% 
  relocate(geometry, .after = environmental_conditions) 

ssp5_pu_costs <- ssp5_pu %>% 
  full_join(costs_cnr_cad, by = c("biome_code")) %>% 
  relocate(geometry, .after = environmental_conditions)  
```




### Features
To evaluate carbon and biodiversity as features in the prioritization problem, they must be merged into a raster stack.

```{r}
# convert biodiversity and carbon layers from
# terra "SpatRaster" format to raster "RasterLayer"
biodiversity_brazil <- raster(biodiversity_brazil)
carbon_brazil <- raster(carbon_brazil)

# now stack the two rasters
bio_carbon <- stack(biodiversity_brazil, carbon_brazil)


```




## Problem Formulations

This is where we will start formulating the problem to be solved in prioritizr!
```{r}
# test problem
ssp1_abandoned_crop <- raster(ssp1_abandoned_crop)
p1 <- problem(ssp1_abandoned_crop, bio_carbon) %>% 
  add_min_set_objective()

# print number of planning units
number_of_planning_units(p1)

# print number of features
number_of_features(p1)

```
 









